---
title: "composerr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{composerr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This **R package** makes it easy to add a
**full range and maintainable error handling**
to your functions and packages, without inflating your code.

> Make error handling in R less tedious


## Installation

```{r}
# Install development version from GitHub
devtools::install_github('a-maldet/composerr', build_opts = NULL)
```

## Usage

```{r}
library(composerr)
```

### Creating advanced error handlers

The call `err_h <- composerr(text_1 = "TEST", sep_1 = ": ")`
creates a `composerr` class object `err_h`.
The created object `err_h` is basically a function `err_h = function(msg)`,
which throws an error with an concatenated error message (`text_1 + sep_1 + msg`)
when called.

The resulting error handling function also has an argument `internal_handler`.
This argument defines the internal processing of the created error message.
Usually `intenal_handler = stop`, but you can define any other internal 
processing of the passed in error message, like:

- throwing an error (**default**): `internal_handler = message`
- sending a message: `internal_handler = message`
- sending a warning: `internal_handler = warning`
- writing the message into a log file:
  `internal_handler = function(msg) cat(msg, FILEPATH)`
- or any other custom behavior, that can be achieved with a function
  `function(msg) {...}`

By setting

```{r}
err_h_x <- composerr("MORE DETAILS", err_h)
```
one
can create a new error handler `err_h_x()` from the previously created
error handler `err_h()`. When calling 
```{r}
err_h_x("PROBLEM DESCRIPTION")
```
then `err_h()` is called by `err_h_x()`, but first the additional message part
of `"MORE DETAILS"` is added to `"PROBLEM DESCRIPTION"`.
Therefore, `err_h_x()` is a **more specific**
error handler than the original error handler `err_h()`.
Overwriting 
```{r}
err_h <- composerr("MORE DETAILS", err_h)
```
is also possible and sometimes useful.

The following **example** shows the implementation of a vector multiplication
function **with complete error handling**:
```{r}
my_vec_mult <- function(x, y) {
  # create your error handlers
  err_h <- composerr("In `my_vec_mult()`")
  err_h_x <- composerr("Invalid argument `x`", err_h)
  err_h_x <- composerr("Invalid argument `y`", err_h)
  if (!is.numeric(x))
    err_h_x("Not a number.")
  if (any(is.na(x)))
    err_h_x("Has missings.", internal_handler = warning)
  if (!is.numeric(y))
    err_h_x("Not a number.")
  if (any(is.na(y)))
    err_h_x("Has missings.", internal_handler = warning)
  if (length(x) != length(y))
    err_h("Vectors `x` and `y` have different length.")
  sum(x*y, na.rm = TRUE)
}
my_vec_mult("a", 1:2)
# Error: In `my_vec_mult()`: Invalid argument `x`: Not a number.
my_vec_mult(c(1, NA), 1:2)
# Warning: In `my_vec_mult()`: Invalid argument `x`: Has missings.
# 1
my_vec_mult(1:2, "b")
# Error: In `my_vec_mult()`: Invalid argument `y`: Not a number.
my_vec_mult(1:2, c(1, NA))
# Warning: In `my_vec_mult()`: Invalid argument `y`: Has non-finite values.
# 1
my_vec_mult(1:2, 1:3)
# Error: In `my_vec_mult()`: Vectors `x` and `y` have different length.
my_vec_mult(1:2, 1:2)
# 14
```

### Halt error processing and flush entire error stack at once

Sometimes it is useful to halt the error processing and just
collect the error messages in an internal error stack and
flush the entire stack later on.
This can be done by calling:

- `composerr_halt(err_h)`: Halt the error processing
- `composerr_flush(err_h)`: Flush the entire error stack at once

The following **example** shows a more advanced implementation of `my_vec_mult2()`
by using a validation routine with a **complete error handling**:
```{r}
validate_numeric_vec <- function(obj, err_h) {
  obj_name <- deparse(substitute(obj))
  err_h <- composerr(paste0("Invalid argument `", obj_name, "`"), err_h)
  if (!is.numeric(obj))
    err_h("Not a number.")
  err_h <- composerr(err_h = err_h)
  composerr_halt(err_h)
  for (i in seq_along(obj)) {
    err_h_item <- composerr(text_1 = paste0("Item-", i), err_h, sep_1 = " is ")
    if (is.na(obj[i]) && !is.nan(obj[i]))
      err_h_item("NA.")
    if (is.nan(obj[i]))
      err_h_item("NaN.")
    if (is.infinite(obj[i]))
      err_h_item("infinite.")
  }
  composerr_flush(err_h)
  invisible(obj)
}
my_vec_mult2 <- function(x, y) {
  err_h <- composerr("In `my_vec_mult2()`")
  validate_numeric_vec(x, err_h)
  validate_numeric_vec(y, err_h)
  if (length(x) != length(y))
    err_h("Vectors `x` and `y` have different length.")
  sum(x*y)
}
my_vec_mult2("a", 1:4)
# Error: In `my_vec_mult2()`: Invalid argument `x`: Not a number.
my_vec_mult2(c(1, NA, NaN, Inf, 5), 1:5)
# Error: In `my_vec_mult2()`: Invalid argument `x`:
#   - Item-2 is NA.
#   - Item-3 is NaN.
#   - Item-4 is infinite.
my_vec_mult2(1:5, c(NaN, 2, 3, NA, Inf))
# Error: In `my_vec_mult2()`: Invalid argument `y`:
#   - Item-1 is NA.
#   - Item-4 is NaN.
#   - Item-5 is infinite.
my_vec_mult2(1:5, 1:4)
# Error: In `my_vec_mult2()`: Vectors `x` and `y` have different length.
my_vec_mult2(1:5, 1:5)
55
```

### List of contained functions

- `composerr()`: Create new error handling function from scratch.
- `composerr(err_h = err_h)`: Create new error handling function from the
  previously created error handler `err_h()`.
- `composerr_modify(err_h)`: Modify the existing error handling function
  `err_h()` without creating a new one.
- `composerr_halt(err_h)`: Halt error processing and accumulate all error 
  messages generated by calling `err_h()` in an error stack.
- `composerr_flush(err_h)`: Flush entire error stack, holding all stacked
  error messages generated by calling `err_h()`.
- `composerr_counterr(err_h)`: Number of accumulated error messages
  generated by calling `err_h()`.
- `composerr_get_internal_handler(err_h)`: Retrieve the internal error
  processing function of `err_h()`.
- `composerr_validate(err_h)`: Validate that `err_h()` is indeed an error
  handler created by `composerr()`.
  
## Remark:

If you like this package, please give me a star on github:
[https://github.com/a-maldet/composerr](https://github.com/a-maldet/composerr)

## License

GPL-3
